---
title: "Nested Modelling for Suffixes"
author: Eva, Utku, Özge
format:
  html:
    self-contained: true  # ensures everything is embedded, no extra folders
    embed-resources: true
    fig-width: 7
    fig-height: 4
    smooth-scroll: true
    code-copy: true
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    toc: true             # adds table of contents
    toc-location: left    # optional: move TOC to left side
    toc-depth: 3          # optional: control how deep the TOC goes
    number-sections: true # adds numbering to sections
    theme: flatly          # or use another Bootstrap theme like journal, flatly, etc.
    highlight-style: a11y # accessible syntax highlighting
    # css: styles.css       # optional: your custom CSS if needed
editor_options: 
  chunk_output_type: console
---


## Setup: load packages and scripts and data


```{r, warning=FALSE, message=FALSE}
# packages
invisible(lapply(c("dplyr", "ggplot2", "tidyr", "readr", "purrr", "stringr", "knitr", "brms", "tidybayes", "loo"), library, character.only = TRUE))


# load scripts
invisible(lapply(list.files("./scripts", "\\.R$", full.names = TRUE), source))

# load data
exp <- read_csv("./data/exp_data.csv") 

freq <- read_csv("../freqs_for_analysis.csv")

exp_joined <- exp %>%
  rowwise() %>%
  mutate(matched = list(
    freq %>% filter(str_detect(item, str_c("\\b", conj1, "\\b"))))
  ) %>%
  unnest(matched)

exp_joined <- exp_joined %>%
  mutate(logCoord = log(Coord + 1))  %>%
  mutate(logCoord_z = scale(logCoord)[, 1])
```



## Models with Coordination Frequency

```{r}
exp_joined$is_sa <- as.factor(exp_joined$is_sa)
exp_joined$rating <- factor(exp_joined$rating, ordered = TRUE)
exp_joined <- exp_joined %>%
  group_by(completionCode) %>%
  mutate(subject = paste0("S", cur_group_id())) %>%
  ungroup()
exp_joined$subject <- as.factor(exp_joined$subject)
exp_joined$itemNum <- as.factor(exp_joined$itemNum)
exp_joined$suffix <- as.factor(exp_joined$suffix)

contrasts(exp_joined$is_sa) <- -1 * contr.sum(2)/2  # centered contrast for suspension (.5 = Suspended, -.5 = Not Suspended)
contrasts(exp_joined$suffix) <- contr.sum(4)/2 # deviation coding for suffixes

my_prior <- c(
  prior(normal(0, 0.5), class = "b", coef = "is_sa1"),
  prior(exponential(1), class = "sd", group = "subject"),
  prior(exponential(1), class = "sd", group = "itemNum"),
  prior(exponential(1), class = "sd", coef = "is_sa1", group = "subject"),
  prior(lkj(2), class = "cor"),
  prior(normal(0, 5), class = "Intercept")
)
```


```{r}

run_suffix_model <- function(suffix_level, data, prior, model_path_prefix = "./models/", freq_var = "logCoord_z") {
  # Create formula dynamically
  formula_str <- paste0("rating ~ 1 + is_sa * ", freq_var, 
                        " + (1 + is_sa * ", freq_var, " | subject) + (1 + is_sa | itemNum)")
  
  # Fit the model
  fit <- brm(
    formula = bf(as.formula(formula_str), decomp = "QR"),
    data = data %>% filter(suffix == suffix_level),
    family = cumulative(link = "logit"),
    prior = prior,
    chains = 4, iter = 4000, cores = 8,
    control = list(adapt_delta = 0.9, max_treedepth = 16),
    file = paste0(model_path_prefix, "model_", gsub("-", "", suffix_level), "_", freq_var)
  )

  # Extract fixed effects
  interaction_term <- paste0("b_is_sa1:", freq_var)
  posterior_df <- fit %>%
    spread_draws(!!sym("b_is_sa1"), !!sym(paste0("b_", freq_var)), !!sym(interaction_term)) %>%
    select(-.chain, -.iteration, -.draw) %>%
    pivot_longer(cols = everything(), names_to = "term", values_to = "estimate")

  # Summarise with posterior probabilities and CIs
  posterior_summary <- posterior_df %>%
    group_by(term) %>%
    summarise(
      p_gt_0 = mean(estimate > 0),
      avg_estimate = mean(estimate),
      ci_95_low = quantile(estimate, 0.025),
      ci_95_high = quantile(estimate, 0.975),
      .groups = "drop"
    )

  # Join summary for plotting
  posterior_df <- left_join(posterior_df, posterior_summary, by = "term")
  x_range <- range(posterior_df$estimate)
  buffer <- diff(x_range) * 0.3
  max_x <- max(posterior_df$estimate)

  # Plot
  p <- ggplot(posterior_df, aes(x = estimate, y = reorder(term, avg_estimate))) +
    stat_halfeye(.width = c(.66, .95)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    geom_text(
      data = posterior_summary,
      aes(
        x = max_x + buffer * 0.1,
        y = term,
        label = paste0("P(β>0) = ", round(p_gt_0, 2))
      ),
      hjust = 0,
      inherit.aes = FALSE
    ) +
    labs(
      x = "Effect (log odds)",
      y = "Coefficient",
      title = paste("Posterior Distributions for", suffix_level, "with", freq_var)
    ) +
    theme_minimal() +
    coord_cartesian(xlim = c(x_range[1], max_x + buffer))

  return(list(model = fit, plot = p, summary = posterior_summary))
}

```

### -lI

- No main effect of suspension
- Coordination frequency overall reduces ratings
- Negative Interaction: higher coordination frequency leads to lower ratings for suspended suffixes

```{r}
result_lI <- run_suffix_model("-lI", exp_joined, my_prior)
result_lI$summary
result_lI$plot
```

### -lIk

- No main effect of frequency
- Almost no main effect of suspension
- No interaction

```{r}
result_lIk <- run_suffix_model("-lIk", exp_joined, my_prior)
result_lIk$summary
result_lIk$plot
```

### -sIz

- No main effect of frequency
- Suspension reduces overall ratings
- Positive interaction between suspension and frequency: higher coordination frequency leads to higher ratings for suspended suffixes
```{r}
result_sIz <- run_suffix_model("-sIz", exp_joined, my_prior)
result_sIz$summary
result_sIz$plot
```

### -cI


- Frequency overall increases ratings
- Suspension overall reduces ratings
- Positive interaction between suspension and frequency: higher coordination frequency leads to higher ratings for suspended suffixes

```{r}
result_cI <- run_suffix_model("-cI", exp_joined, my_prior)
result_cI$summary
result_cI$plot
```


## Models with csumPMI Frequency

```{r}
exp_joined <- exp_joined %>%
  mutate(log_csumPMI = log1p(csumPMI)) %>% 
  mutate(log_csumPMI_z = scale(log1p(csumPMI)))
```


### -lI
 
- No effect

```{r}
result_lI <- run_suffix_model("-lI", exp_joined, my_prior, freq_var = "log_csumPMI_z")
result_lI$summary
result_lI$plot
```

### -lIk

- No effect

```{r}
result_lIk <- run_suffix_model("-lIk", exp_joined, my_prior,freq_var = "log_csumPMI_z")
result_lIk$summary
result_lIk$plot
```

### -sIz

- Main negative effect of suspended affixation
- Main positive effect of frequency
- No interaction

```{r}
result_sIz <- run_suffix_model("-sIz", exp_joined, my_prior,freq_var = "log_csumPMI_z")
result_sIz$summary
result_sIz$plot
```

### -cI

- Main positive effect of frequency
- Main negative effect of suspended affixation
- Maybe a positive interaction? But very unlikely.

```{r}
result_cI <- run_suffix_model("-cI", exp_joined, my_prior,freq_var = "log_csumPMI_z")
result_cI$summary
result_cI$plot
```

